---
title: "ReproducibleResearchPeer1"
author: "MSP"
date: "September 23, 2016"
output: html_document
---

## Setup

```{r setup, include=FALSE}
        knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8, fig.path='Figs/')
        library(data.table)
        library(ggplot2)
        library(dplyr)
```

## Load and preprocess the data

```{r Read Data}
        # set global project directory
        mainDir <- "C:/ReproducibleResearchPeer1"
        filename <- "activity.csv"
        setwd(mainDir)
        amd <- read.csv(file=filename, head=TRUE, sep=",")
```

## What is mean total number of steps taken per day?

```{r Calculate Total Steps, echo=FALSE}
        # calculate mean steps per day
        setDT(amd)[, TotalStepsDay := sum(steps, na.rm = TRUE), by = date]
        # calculate bins
        b<-length(unique(amd$date))
        # plot total steps per day
        qplot(amd$TotalStepsDay, 
              geom="histogram",
              bins=b,
              main = "Total Steps Per Day", 
              xlab = "Total Steps",  
              fill=I("blue"), 
              col=I("red"))
```

## Mean total number of steps taken per day

```{r Calculate Total Steps mean, echo=FALSE}
        mean(amd$TotalStepsDay)
```

## Median total number of steps taken per day

```{r Calculate Total Steps median, echo=FALSE}
        median(amd$TotalStepsDay)
```

## What is the average daily activity pattern?

```{r Plot steps average, echo=FALSE}
        # calculate mean steps per interval across all days
        setDT(amd)[, AvgStepsInterval := mean(steps, na.rm = TRUE), by = interval]
        # plot mean steps per interval across all days
        ggplot(amd, 
          aes(interval, AvgStepsInterval)) + 
          geom_line() +
          xlab("Interval") + 
          ylab("Avg Steps")
```

## Max Interval Value

```{r Max interval, echo=FALSE}
        # show max interval
        amd[which.max(amd$AvgSteps),]$interval
```

## Impute missing values

```{r Missing Values, echo=FALSE}
        # impute missing steps using average for interval
        amd2 <- amd %>% mutate(isteps = ifelse(is.na(steps),as.integer(round(AvgStepsInterval,digits=0)),steps))
        # calculate mean steps per day
        setDT(amd2)[, ITotalStepsDay := sum(isteps), by = date]
        # calculate bins
        b<-length(unique(amd2$date))
        # plot total steps per day
        qplot(amd2$ITotalStepsDay, 
              geom="histogram",
              bins=b,
              main = "Total Steps Per Day Imputed", 
              xlab = "Total Steps Imputed",  
              fill=I("blue"), 
              col=I("red"))
```

## Rows with missing values

```{r Missing Values Rows, echo=FALSE}
        # show numbers of rows with missing values for steps
        sum(is.na(amd$steps))
```

## Mean total number of steps taken per day with missing values imputed

```{r With Missing Values Mean, echo=FALSE}
        mean(amd2$ITotalStepsDay)
```

## Median total number of steps taken per day with missing values imputed

```{r With Missing Values Median, echo=FALSE}
        median(amd2$ITotalStepsDay)
```

## Weekdays versus weekends

```{r Weekdays and Weekends, echo=FALSE}
        amd3 <- amd2 %>% mutate(DayType = as.factor(
                ifelse(weekdays(as.Date(as.character(amd2$date))) %in% c('Sunday','Saturday'),"weekend","weekday")))
        # calculate mean steps per interval across Weekends or WeekDays
        setDT(amd3)[amd3$DayType %in% c('weekend'), AvgStepsWeeks := mean(isteps), by = interval]
        setDT(amd3)[amd3$DayType %in% c('weekday'), AvgStepsWeeks := mean(isteps), by = interval]
        par(mfrow=c(2,1)) 
        ggplot(amd3[amd3$DayType %in% c('weekend'),], 
          aes(interval, AvgStepsWeeks)) + 
          ggtitle("Weekend") +
          geom_line() +
          xlab("Interval") + 
          ylab("Number of Steps")
        ggplot(amd3[amd3$DayType %in% c('weekday'),], 
          aes(interval, AvgStepsWeeks)) + 
          ggtitle("Weekday") +
          geom_line() +
          xlab("Interval") + 
          ylab("Number of Steps")
```
